\documentclass[12pt]{report}
\input{../preamble/preamble}
\addbibresource{../bibliography.bib}

\begin{document}

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}  % Add an entry to the ToC.

First and foremost, I must thank my supervisor Rich.
Rich, you are one of the sharpest, most able, and most knowledgeable persons I know.
It has been a privilege to pursue a PhD under your supervision.

I have only reached the end smiling
because great friends make
Cambridge feel like home.
Eric and Coz, thank you for your friendship and
innumerable great times.
Our nights in '90s will be remembered.
Will and Joel, I really value hanging out together.
Together with Phil, the legacy of 308 lives on.

My time in Cambridge has been incredibly inspiring.
Andrew, David,
James, Stratis,
Ji\v{r}\'{i},
Jonathan,
Anna,
and Tom,
I thoroughly enjoyed working together.
Thank you for many insightful discussions.
I have learned a great amount from every one of you.
I thank Invenia for the support throughout the years,
and all Invenians for interesting and fruitful discussions.

Back home, I also enjoyed the support of great friends.
Pep en Oscar, thank you for countless many good nights.
May there be many more to come.
\textit{Adt fundum!}
Bram,
Yann, and Mike,
thank you for your long-standing friendships.
Peter en Manouk, our new buurtjes,
I look forward to spending more time together.

Coming home after having been away has always been a source of comfort.
Pap, mam, and Merel, thank you for everything.
Opa,
you would have loved to see this thesis.
I hope this thesis makes you proud.

Finally, Liesje, my dearest, thank you for your love and endless support.

\end{document}
