\documentclass[12pt]{report}
\input{../../preamble/preamble}
\addbibresource{../../bibliography.bib}

\begin{document}

\chapter{\texorpdfstring{$L^p$-$\beta$}{Lp-beta}-H\"older Equicontinuity}
\label{chap:holder}

\section{Weak Convergence}
Let $(\mu_i)_{i=1}^\infty \sub \mathcal{P}$ and $\mu \in \mathcal{P}$.
We say that $\mu_i$ converges weakly to $\mu$ and write $\mu_i \weakto \mu$ if $\mu_i(L) \to \mu(L)$ for all continuous bounded functions $L\colon C(\mathcal{X}, \mathcal{Y}) \to \R$.
Consider the natural embedding $\iota_R \colon C(\mathcal{X}, \mathcal{Y}) \to C([-R, R], \mathcal{Y})$ and denote $\mu|_{[-R, R]} = \iota_R(\mu)$.
The following proposition allows us to deduce weak convergence from weak convergence on all compacts.

\begin{proposition} \label{prop:weak_convergence_from_compacts}
    $\mu_i \weakto \mu$ if and only if $\mu_i|_{[-R,R]} \weakto \mu|_{[-R, R]}$ for all $R > 0$.
\end{proposition}

\begin{proof}
    Let $L\colon C(\mathcal{X}, \mathcal{Y}) \to \R$ be continuous and bounded.
    Let $R > 0$.
    Consider the particular right inverse $\iota_R^\dagger \colon C([-R, R], \mathcal{Y}) \to C(\mathcal{X}, \mathcal{Y})$ of $\iota_R$ defined by $\iota_R^\dagger(f)(x) = f(\sign(x)(\abs{x} \land R))$.
    By construction of $d_C$, we can find $\e > 0$ such that, for all $f, g \in C(\mathcal{X}, \mathcal{Y})$, we have $\abs{L(f) - L(g)} < \e$ whenever $g|_{[-R, R]} = f|_{[-R, R]}$.
    Therefore,
    \begin{equation}
        \sup_{i \ge 1}\,
        \abs{
            \E_{\mu_i\cond_{[-R, R]}}[L \comp \iota_R^\dagger(f)] -
            \E_{\mu_i}[L(f)]
        }
        \le
        \sup_{i \ge 1}\,
        \E_{\mu_i}[\abs{
            L(\iota_R^\dagger \comp \iota_R(f)) -
            L(f)
        }]
        \le \e,
    \end{equation}
    because $\iota_R^\dagger \comp \iota_R(f)|_{[-R, R]} = f_{[-R, R]}$.
    In other words,
    \begin{equation}
        \lim_{R \to \infty} \E_{\mu_i\cond_{[-R, R]}}[L \comp \iota_R^\dagger(f)] = \E_{\mu_i}[L(f)] \quad \text{uniformly over $i \ge 1$}.
    \end{equation}
    Then compute
    \begin{align*}
        \lim_{i \to \infty} \E_{\mu_i}[L(f)]
        &= \lim_{i \to \infty} \lim_{R \to \infty} \E_{\mu_i|_{[-R, R]}}[L \comp \iota_R^\dagger(f)] \\
        &=  \lim_{R \to \infty} \lim_{i \to \infty} \E_{\mu_i|_{[-R, R]}}[L \comp \iota_R^\dagger(f)] \\
        &=  \lim_{R \to \infty} \E_{\mu|{[-R, R]}}[L \comp \iota_R^\dagger(f)] \\
        &= \E_{\mu}[L(f)]
    \end{align*}
    where the interchange of limits is justified by the aforementioned uniformity of convergence.
    Since $L$ was arbitrary, this shows that $\mu_i \weakto \mu$.
\end{proof}

\section{\texorpdfstring{$L^{p,q}$}{Lpq}-H\"older Equicontinuity}

We will want to deduce weak convergence from weak convergence of the finite-dimensional distributions.
For this, we define the notion of \emph{$L^{p,q}$-H\"older equicontinuity}.
Let $\mu \in \mathcal{P}$.
Call a single measure $\mu$ \emph{$L^{p,q}$-H\"older continuous} for some $p > 0$ and $q > 0$ if there exist a constant $c > 0$ and a radius $r > 0$ such that
\begin{equation}
    \E_\mu[\abs{f(x) - f(y)}^p] \le c \abs{x - y}^q
    \quad\text{whenever}\quad
    \abs{x - y} < r.
\end{equation}
Let $\mathcal{F} \sub \mathcal{P}$ be a family of probability measures.
Call this family \emph{$L^{p,q}$-H\"older equicontinuous} if there exist a constant $c > 0$ and a radius $r > 0$ such that
\begin{equation}
    \sup_{\mu \in \mathcal{F}}\,
    \E_\mu[\abs{f(x) - f(y)}^p] \le c \abs{x - y}^q
    \quad\text{whenever}\quad
    \abs{x - y} < r.
\end{equation}
This can also be written as a limit:
\begin{equation}
    \lim_{h \to 0}\,
    \sup_{\mu \in \mathcal{F}}\,
    \sup_{x \in \X}\,
    \frac{\E_\mu[\abs{f(x + h) - f(x)}^p]}{h^q}
    < \infty.
\end{equation}
If $\F$ are Gaussian processes, then the criterion simplifies.

\begin{proposition} \label{prop:Holder_continuity_Gaussians}
    Let $\F\ss{G} \sub \mathcal{P}\ss{G}$ be a family of non-degenerate Gaussian processes.
    For $\mu \in \F\ss{G}$, denote its mean function by $m_\mu\colon \mathcal{X} \to \R$ and its covariance function by $k_\mu\colon \X \times \X \to \R$.
    Let $p \ge 2$.
    Then $\F\ss{G}$ is $L^{p, q}$-H\"older equicontinuous if \note{[and only if?]} it is $L^{2, 2q/p}$-H\"older equicontinuous:
    \begin{align*}
        &\lim_{h \to 0}\,
        \sup_{\mu \in \F}\,
        \sup_{x \in \X}\,
        \frac{(m_\mu(x + h) - m_\mu(x))^2 + \Delta k_\mu(x + h, x)}{h^{2q/p}} \\
        &\qquad =
        \lim_{h \to 0}\,
        \sup_{\mu \in \F}\,
        \sup_{x \in \X}\,
        \frac{\E_\mu[\abs{f(x + h) - f(x)}^2]}{h^{2q/p}}
        < \infty
    \end{align*}
    where
    \begin{equation}
        \Delta k_\mu(x + h, x)
        = k_\mu(x + h, x + h) + k_\mu(x, x) - 2 k_\mu(x + h, x).
    \end{equation}
    Consequently, if $\F\ss{G}$ is $L^{p,q}$-H\"older equicontinuous for $p \ge 2$ if and only if $\F\ss{G}$ is $L^{\alpha p,\alpha q}$-H\"older equicontinuous with $\alpha > 0$ for $\alpha p \ge 2$.
\end{proposition}
\begin{proof}
    Note that
    \begin{equation}
        f(x + h) - f(x)
        \sim \Normal(m_\mu(x + h) - m_\mu(x), \Delta k_\mu(x + h, x)).
    \end{equation}
    For any $U \sim \Normal(\mu, \sigma^2)$, it holds that
    \begin{equation}
        \E[\abs{U}^p]
        \le 2^{p} M_p (\abs{\mu}^p + \sigma^{p})
    \end{equation}
    where $M_p = \E[\abs{N}^p] < \infty$ with $N \sim \Normal(0, 1)$.
    Therefore,
    \begin{equation}
        \frac{\E_\mu[\abs{f(x + h) - f(x)}^p]}{h^q}
        \le
        2^p M_p \frac{((m_\mu(x + h) - m_\mu(x))^2)^{p/2} + (\Delta k_\mu(x + h, x))^{p/2}}{h^q}.
    \end{equation}
    Since $p \ge 2$, $a^{p/2} + b^{p/2} \le (a + b)^{p/2}$ for $a, b \ge 0$, so
    \begin{equation}
        \frac{\E_\mu[\abs{f(x + h) - f(x)}^p]}{h^q}
        \le
        2 M_p\frac{ ((m_\mu(x + h) - m_\mu(x))^2 + \Delta k_\mu(x + h, x))^{p/2}}{h^q}.
    \end{equation}
    Hence, it is sufficient that
    \begin{equation}
        \lim_{h \to 0}\,
        \sup_{\mu \in \F}\,
        \sup_{x \in \X}\,
        2^p M_p
        \frac{((m_\mu(x + h) - m_\mu(x))^2 + \Delta k_\mu(x + h, x))^{p/2}}{h^q}
        < \infty.
    \end{equation}
    Equivalently, it is sufficient that
    \begin{equation}
        \lim_{h \to 0}\,
        \sup_{\mu \in \F}\,
        \sup_{x \in \X}\,
        \frac{(m_\mu(x + h) - m_\mu(x))^2 + \Delta k_\mu(x + h, x)}{h^{2q/p}}
        < \infty.
    \end{equation}
    Here the numerator is equal to $\E_\mu[\abs{f(x + h) - f(x)}^2]$, so we can also write
    \begin{equation}
        \lim_{h \to 0}\,
        \sup_{\mu \in \F}\,
        \sup_{x \in \X}\,
        \frac{\E_\mu[\abs{f(x + h) - f(x)}^2]}{h^{2q/p}}
        < \infty,
    \end{equation}
    which concludes the proof.
\end{proof}

If the Gaussian processes in $\F\ss{G}$ are stationary, then the condition simplifies to
\begin{equation}
    \lim_{h \to 0}\,
    \sup_{\mu \in \F\ss{G}}\,
    \frac{
        k_\mu(0)
        - k_\mu(h)
    }{h^{2q/p}}
    = 0.
\end{equation}
If the kernels are once differentiable at the origin, then, upon choosing $p = 4$ and $1 < q < 2$, it suffices that $\sup_{\mu \in \F\ss{G}} \norm{k'_\mu}_\infty < \infty$.
For kernels of the form $k_\mu(h) = \exp(-g_\mu(h))$ with $g(0) = 0$ where $g \colon [0, \infty) \to [0, \infty)$ is strictly increasing, we require \note{[double check this!]}
\begin{equation}
    \lim_{h \to 0}\,
    \sup_{\mu \in \F\ss{G}}\,
    \frac{
        g_\mu(h)
    }{h^{2q/p}}
    = 0.
\end{equation}
For example, if $g_\mu(h) = c_\mu h$ for $c_\mu > 0$, then, upon choosing $p = 4$ and $1 < q < 2$, it suffices that $\sup_{\mu \in \F\ss{G}}\, c_\mu < \infty$.

Let $I = \union_{n=1}^\infty \mathcal{X}^n$ denote the collection of all finite index sets.
For $\vx \in \R^n$, let $P_\vx\colon C(\mathcal{X}, \mathcal{Y}) \to \mathcal{Y}^n$ be the projection onto the index set $\vx$: $P_\vx f = (f(x_1), \ldots, f(x_n))$.
For $\vx \in \R^n$ and a measure $\mu \in \mathcal{P}$, let $P_\vx \mu$ be the corresponding pushforward measure.

\begin{lemma} \label{lem:modulus_continuity}
    Suppose that $(\mu_i)_{i=1}^\infty \sub \mathcal{P}$ are H\"older $L^{p,q}$-equicontinuous for some $p > 0$ and $q > 1$;
    that is, there exist $p > 0$, $q > 1$, a constant $c > 0$, and a radius $r > 0$ such that
    \begin{equation}
        \sup_{i \ge 1}\,
        \E_{\mu_i}[\abs{f(x) - f(y)}^p] \le c \abs{x - y}^q
        \quad\text{whenever}\quad
        \abs{x - y} < r.
    \end{equation}
    For $f \in C([0, 1], \Y)$, define the \emph{modulus of continuity} by
    \begin{equation}
        \omega_f(h)
        = \sup_{x, y \in [0, 1]: \abs{x - y} < h} \abs{f(x) - f(y)}.
    \end{equation}
    Then, for all $\e > 0$,
    \begin{equation}
        \lim_{h \to 0} \sup_{i \ge 1}\, \mu_i(\omega_f(h) \ge \e) = 0.
    \end{equation}
\end{lemma}
\begin{proof}
    Choose $k \in \N$ such that $2^{-(k + 1)} \le h \le 2^{-k}$.
    The proof strategy mimics the proof of Theorem 4.2.1 by \textcite{Norris:2018:Advanced_Probability}.
    Let $i \in \N$.
    For $n \in \N$, denote $\Db_n = \set{0, 2^{-n}, 2 \cdot 2^{-n}, \ldots, 1}$.
    Consider $f \in C([0, 1], \mathcal{Y})$.
    Set
    \begin{equation}
        K_n = \sup_{x \in \Db_n \setminus \set{1}} = \abs{f(x + 2^{-n}) - f(x)}.
    \end{equation}
    Overestimate the supremum by a sum and use $L^{p,q}$-H\"older continuity of $\mu_i$:
    \begin{equation}
         \E_{\mu_i}[K_n^p]
         \le \sum_{i=0}^{2^n - 1} \E_{\mu_i}[\abs{f(x + 2^{-n}) - f(x)}^p]
         \le \sum_{i=0}^{2^n - 1} 2^{-q n}
         = 2^{-(q - 1) n}.
    \end{equation}
    For any $x, y \in \union_{n=1}^\infty \Db_n$ such that $x < y < x + 2^{-k}$, note that the interval $[x, y)$ is the finite, disjoint union of intervals $[r, r + 2^{-n})$ with $r \in \Db_n$ for $n \ge k + 1$ where no three intervals have the same length.
    Therefore, for such $x$ and $y$, by continuity of $f$,
    \begin{equation}
        \omega_f(2^{-k})
        = \sup_{x, y \in \union_{n=1}^\infty \Db_n: \abs{x - y}< 2^{-k}}\abs{f(x) - f(y)}
        \le 2\sum_{n=k+1}^\infty K_n.
    \end{equation}
    Hence,
    \begin{equation}
        \E_{\mu_i}\sbrac{\omega_f(2^{-k})}
        % \le 2\sum_{n=k+1}^\infty \E_{\mu_i}[K_n]
        \le 2\sum_{n=k+1}^\infty \E_{\mu_i}[K_n^p]^{\frac1p}
        \le 2\sum_{n=k+1}^\infty 2^{-\frac{q - 1}{p}n}
        = \frac{2^{1-\frac{q - 1}{p}(k + 1)}}{1 - 2^{-\frac{q - 1}{p}}}
        = c_{p,q} 2^{-\frac{q - 1}{p}k},
    \end{equation}
    for some constant $c_{p, q}$ that depends on $p$ and $q$, using that $q > 1$.
    Then, for any $\e > 0$, by Markov's Inequality,
    \begin{equation}
        \sup_{i \ge 1}\, \mu_i(\omega_f(h) \ge \e)
        \le \sup_{i \ge 1}\, \mu_i(\omega_f(2^{-k}) \ge \e)
        \le \frac{c_{p,q}}{\e} 2^{-\frac{q - 1}{p}k}.
    \end{equation}
    This proves the result, because $k \to \infty$ as $h \to 0$.
\end{proof}

\begin{proposition} \label{prop:weak_convergence_from_fdds}
    Let $(\mu_i)_{i=1}^\infty \sub \mathcal{P}$ be weakly convergent to some $\mu \in \mathcal{P}$ along a dense sequence of finite-dimensional distributions: $P_{\vx_k} \mu_i \weakto P_{\vx_k} \mu$ for all $k \ge 1$ where $(\vx_k)_{k \ge 1} \sub I$ is a sequence that becomes dense in $\X$.
    If $(\mu_i)_{i=1}^\infty$ are H\"older $L^{p,q}$-equicontinuous for some $p > 0$ and $q > 1$, then $\mu_i \weakto \mu$.
\end{proposition}
\begin{proof}
    We demonstrate that $\mu_i|_{[-R, R]} \weakto \mu|_{[-R, R]}$ for all $R > 0$ and appeal to \cref{prop:weak_convergence_from_compacts}.
    For notational convenience, we only consider the restriction to $[0, 1]$ and suppress the notation for the restriction.

    For $\vx \in [0, 1]^n$, consider the particular right inverse $P_\vx^\dagger \colon \R^n \to C([0, 1], \mathcal{Y})$ of $P_\vx$ defined by setting $P_\vx^\dagger \vy$ to the linear interpolation through the points $(x_1,y_1), \ldots, (x_n, y_n)$ and letting it be constant outside the extremal inputs.
    Denote $\norm{f - g}_\infty = \sup_{x\in [0, 1]} \abs{f(x) - g(x)}$.

    The restriction $(\vx_k)_{k = 1}^\infty \sub \union_{n=1}^\infty [0, 1]^n$ becomes dense in $[0, 1]$.
    Therefore, by \cref{lem:modulus_continuity}, for every $\delta > 0$,
    \begin{equation}
        \lim_{k \to \infty} \sup_{i \ge 1}\, \mu_i(\norm{P_{\vx_k}^\dagger P_{\vx_k} f - f}_\infty \ge \delta) = 0.
    \end{equation}
    Note that we also have
    \begin{equation}
        \lim_{k \to \infty} \mu(\norm{P_{\vx_k}^\dagger P_{\vx_k} f - f}_\infty \ge \delta) = 0.
    \end{equation}
    This is because, $\mu$--almost surely, $f$ is continuous on a compact set, hence uniformly continuous, so $\norm{P_{\vx_k}^\dagger P_{\vx_k} f - f}_\infty \to 0$ almost surely and consequently in probability.

    Let $L\colon C([0, 1], \mathcal{Y}) \to \R$ be continuous and bounded.
    Let $\e > 0$.
    By continuity of $L$, there exists a $\delta > 0$ such that $\norm{f - g}_\infty < \delta$ implies that $\norm{L(f) - L(g)}_\infty < \e$.
    Consider the event $A^k_\delta = \set{\norm{P_{\vx_k}^\dagger P_{\vx_k} f - f} < \delta}$.
    Then estimate
    \begin{equation}
        \abs{
            \E_{\mu_i}[L(P_{\vx_k}^\dagger P_{\vx_k} f)]
            - \E_{\mu_i}[L(f)]
        }
        \le \e + \norm{L}_\infty \mu_i((A^k_\delta)^\c)
        \le \e + \norm{L}_\infty \sup_{i \ge 1} \,\mu_i((A^k_\delta)^\c).
    \end{equation}
    Therefore, taking $k \to \infty$ shows that
    \begin{equation}
        \lim_{k \to \infty} \E_{P_{\vx_k} \mu_i}[L \comp P_{\vx_k}^\dagger]
        = \E_{\mu_i}[L] \quad \text{uniformly over $i \ge 1$}.
    \end{equation}
    Moreover, the same limit holds for $\mu$.
    Then compute
    \begin{align*}
        \lim_{i \to \infty} \E_{\mu_i}[L]
        &= \lim_{i \to \infty} \lim_{k \to \infty} \E_{P_{\vx_k} \mu_i}[L \comp P_{\vx_k}^\dagger] \\
        &=  \lim_{k \to \infty} \lim_{i \to \infty} \E_{P_{\vx_k}\mu_i}[L \comp P_{\vx_k}^\dagger] \\
        &=  \lim_{k \to \infty} \E_{P_{\vx_k}\mu}[L \comp P_{\vx_k}^\dagger] \\
        &= \E_{\mu}[L]
    \end{align*}
    where the interchange of limits is justified by the aforementioned uniformity of convergence.
    Since $L$ was arbitrary, this shows that $\mu_i \weakto \mu$.
\end{proof}

A lemma that we will need later says that, if $(\mu_i)_{i=1}^\infty \sub \mathcal{P}$ are H\"older $L^{p,q}$-equicontinuous for some $p > 0$ and $q > 1$, then finite-dimensional distributions converge weakly in a uniform way if their inputs converge.

\begin{lemma} \label{lem:Holder_uniform_weak_convergence-fdds}
    Suppose that $(\mu_i)_{i=1}^\infty \sub \mathcal{P}$ are H\"older $L^{p,q}$-equicontinuous for some $p > 0$ and $q > 1$.
    For some $n \in \N$, let $(\vx_k)_{k \ge 1} \sub [0, 1]^n$ be convergent to some $\vx \in [0, 1]^n$.
    Then, for any $L\colon \R^n \to \R$ continuous and bounded,
    \begin{equation}
        \lim_{k \to \infty} \sup_{i \ge 1}\, \abs{
            \E_{P_{\vx_k} \mu_i}[L] - \E_{P_\vx \mu_i}[L]
        }
        = 0.
    \end{equation}
\end{lemma}
\begin{proof}
    Let $\e > 0$.
    Then there exists a $\delta > 0$ such that $\abs{L(\vy) - L(\vy')} < \e$ whenever $\norm{\vy - \vy'}_\infty < \delta$.
    Consider the event $A_\delta^k = \set{\norm{P_{\vx_k} f - P_{\vx} f}_\infty < \delta}$.
    Then estimate
    \begin{equation}
        \abs{
            \E_{\mu_i}[L \comp P_{\vx_k}]
            - \E_{\mu_i}[L \comp P_{\vx}]
        }
        \le
            \e + \norm{L}_\infty \mu_i((A^k_\delta)^\c)
        \le
            \e + \norm{L}_\infty \sup_{i \ge 1}\, \mu_i((A^k_\delta)^\c),
    \end{equation}
    and use \cref{lem:modulus_continuity} to take $k \to \infty$ and conclude.
\end{proof}


\end{document}
